---
title: "Zoono Hand Sanitiser"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
pacman::p_load(dplyr,ggplot2,tidyr,brms,janitor,vroom,hrbrthemes)
```

## Import the data


```{r data import}
df<-vroom::vroom("../Data/gerba.data.20200302.csv") %>% janitor::clean_names() %>% select(-1) %>% 
  mutate(treatment=as.factor(treatment),
         group=as.factor(group))
# df %>% 
#   filter(treatment=="Untreated") %>% 
#   filter(time==0) %>% 
#   mutate(conc=)
```

You need to duplicate each person's untreated hand count and label it as treatment to allow for the same statrting value
```{r}

df %>% 
  filter(treatment=="Untreated") %>% 
  filter(time==0) %>% 
  mutate()
```


## Including Plots

Let's have a quick look at each person's individual trajectory. Note that individuals are in both "treated" and "untreated" groups because only one hand is washed with the other one being thte control

```{r Plot the data - Individual, echo=FALSE}
df %>% 
  ggplot()+
  geom_point(aes(x=time,y=conc,colour=id %>% as.factor()),size=1.5)+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x)
 )+
  facet_wrap(~group,nrow = 2)+
  annotation_logticks(sides = "l")  +
  scale_color_brewer(palette = "Set1")+
  xlab("Hours after cleaning")+
  ylab("Concentration [CFU/cm2]")+
  labs(col="Individual participant")+
  hrbrthemes::theme_ipsum()
  
  
```
Is there a correlation between each participant's left and right hand (ie. treated vs untreated). The plot shows the effect of adding the group into the regression.

```{r individual correlation}
#Don't iclude group in the regression
df %>%
    pivot_wider(names_from = treatment, values_from = conc) %>% 
  ggplot()+
  geom_point(aes(x=Untreated,y=Treated,colour=time %>% as.numeric()),size=1.5)+
  geom_smooth(aes(x=Untreated,y=Treated),method="lm")+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),labels = scales::number_format(accuracy = 1)
 )+
      scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),labels = scales::number_format(accuracy = 1)
 )+
  annotation_logticks(sides = "lb")  +
  # scale_color_brewer(palette = "Set1")+
  xlab("Untreated hand CFU")+
  ylab("Treated hand CFU")+
  labs(col="Time after treatment")+
  hrbrthemes::theme_ipsum()+
  theme(axis.text.x=element_text(angle = 90, hjust = 0),legend.position = "top")->a

#Include group in the regression
df %>%
    pivot_wider(names_from = treatment, values_from = conc) %>% 
  ggplot()+
  geom_point(aes(x=Untreated,y=Treated,colour=group,shape=group),size=1.5, show.legend = F)+
  geom_smooth(aes(x=Untreated,y=Treated,colour=group),method="lm")+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),labels = scales::number_format(accuracy = 1)
 )+
      scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),labels = scales::number_format(accuracy = 1)
 )+
  annotation_logticks(sides = "lb")  +
  scale_color_brewer(palette = "Set1")+
  xlab("Untreated hand CFU")+
  ylab("Treated hand CFU")+
  labs(col="Time after treatment")+
  hrbrthemes::theme_ipsum()+
  theme(axis.text.x=element_text(angle = 90, hjust = 0),legend.position = "top")->b

ggpubr::ggarrange(a,b,common.legend = FALSE,labels = c("Ungrouped data","Split by group"))

```


Linear regression on concentration amongst participants. Is there a significant effect of group on the correlation between participants' treated and untreated hands? We perform step-wise regression until no further variables can be removed.
```{r}
df %>%
    pivot_wider(names_from = treatment, values_from = conc) %>% 
  mutate(id=as.factor(id))->data
lm_fit_aggregated<-lm(log10(Treated)~log10(Untreated),data=data)
lm_fit_disaggregated<-lm(log10(Treated)~log10(Untreated)+group,data=data)
# broom::tidy(lm_fit)
# aov(lm_fit_aggregated,lm_fit_disaggregated)
# anova(lm_fit_aggregated,lm_fit_disaggregated)
data %>% checkmate::anyMissing()

lm_fit<-lm(log10(Treated)~.,data=data )
selectedMod <- step(lm_fit)
broom::tidy(selectedMod)

lm_fit<-lm(log10(conc)~.,data=df %>% select(-id) )
selectedMod <- step(lm_fit)
broom::tidy(selectedMod)


```

It turns out that only time and the concentration on the untreated value have a significant effect on predicting concentration on the hand. This is good to know and shows that the regression lines plotted above are not significantly different.

Here we plot the points and superimpose a meand value plus 95% confidence interval

```{r Plot the data - Aggregate, echo=FALSE}
df %>% 
  ggplot()+
  geom_point(aes(x=time,y=conc),colour="gray",alpha=0.5,size=1.5)+
  stat_summary(aes(x=time,y=conc,colour=treatment),fun.data = "mean_cl_boot", size = 1.01)+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x)
 )+
  annotation_logticks(sides = "l")  +
  scale_color_brewer(palette = "Set1")+
  xlab("Hours after cleaning")+
  ylab("Colony forming units [CFU]")+
  labs(col="Group")+
  hrbrthemes::theme_ipsum()
  
  
```

# Regression on time

The issue is that the data are not really linearly related so we need to account for the non-monotonic behaviour within the model.
```{r}
df %>% 
  ggplot()+
  geom_point(aes(x=time,y=conc),colour="gray",alpha=0.5,size=1.5)+
  stat_summary(aes(x=time,y=conc,colour=treatment),fun.data = "mean_cl_boot", size = 1.01)+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x)
 )+
  annotation_logticks(sides = "l")  +
  scale_color_brewer(palette = "Set1")+
  geom_smooth(method="lm",aes(x=time,y=conc,colour=treatment))+
  xlab("Hours after cleaning")+
  ylab("Colony forming units [CFU]")+
  labs(col="Group")+
  hrbrthemes::theme_ipsum()
```


# Run a regression model to estimate the effect of time
Use the bayesian regression package BRMS and plots using bayesplot "group+time+intervention + (time+intervention | participant)"

## Summary statistics 
This can be useful to produce the prior prediction if we have some idea of an initial distribution.

```{r}
df %>% 
  summarise(m=mean(conc),sd=sd(conc)) 
```

# 09/04/2023

```{r lmer}
library(lme4)

# Create a new variable for the time segments (0-2 hours and after 2 hours)
df$TimeSegment <- ifelse(df$time <= 2, "0-2_hours", "after_2_hours")

# Fit the piecewise linear mixed effects model
model <- lmer(conc ~ time * treatment * TimeSegment + (1 | id), data = df)


```
```{r}
# Calculate AIC and BIC
model_AIC <- AIC(model)
model_BIC <- BIC(model)

print(c(model_AIC,model_BIC))
print(model_BIC)

```



```{r fit the Bayesian model}

# fit_m <- brm(log10(conc) ~ 1+time+treatment+ (time+treatment | id), 
#                               data   = df, 
#                               warmup = 1000, 
#                               iter   = 3000, 
#                               chains = 4, 
#                               inits  = "random",
#                               cores  = 4)
# summary(fit_m)
# 
# plot(fit_m)

#Make sure id is a factor
df %>% 
  mutate(id=as.factor(id))->df

fit_m1<-brm(conc~group+time*treatment*TimeSegment+(1|id),data=df,family =Gamma(link = "log"))

fit_m2<-brm(conc~group+time*treatment+(1|id),data=df,family =lognormal(link = "identity", link_sigma = "log"))
           # prior=prior(normal(66569.89,107527.2)),sample_prior = "only")



# conditional_effects(fit_m1)
summary(fit_m1)
# plot(fit_m1)
```

Let's have a look at the density estimates from the model fit_m1 vs fit_m2

```{r}
ggpubr::ggarrange(pp_check(fit_m1), pp_check(fit_m2))
# plot(fit_m2)
```

We have assumed that the treatment has a gaussian distribution. Light blue Plot the posterior predictive values.

```{r ecfd_overlay }
ggpubr::ggarrange(pp_check(fit_m1,type="ecdf_overlay",ndraws = 50),pp_check(fit_m2,type="ecdf_overlay",ndraws = 50))

```



Using bayesplot package we can create some nice plots which investigate the fitted model

```{r}
color_scheme_set("red")
bayesplot::ppc_dens_overlay(y = (fit_m2$data$conc),
                 yrep = posterior_predict(fit_m2, draws = 5))
```
### Model selection  

Efficient approximate leave-one-out cross-validation (LOO) 

```{r}
#Model with Gamma respones
(loo1 <- loo(fit_m1))


#Model with log-normal response
(loo2 <- loo(fit_m2))   

# compare both models
loo_compare(loo1, loo2) 


```

Loo suggests that the lognormal response is a marginally better fit than the Gamma.

#### Old test
Epilepsy data

```{r}


fit1 <- brm(count ~ zAge + zBase * Trt + (zBase * Trt|patient), 
            data = epilepsy, family = poisson())

plot(fit1, variable = c("b_Trt1", "b_zBase"))
```


