---
title: "Zoono Hand Sanitiser"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
pacman::p_load(dplyr,ggplot2,tidyr,brms,janitor,vroom,hrbrthemes)
```

## Import the data


```{r data import}
df<-vroom::vroom("../Data/gerba.data.20200302.csv") %>% janitor::clean_names() %>% select(-1) %>% 
  mutate(treatment=as.factor(treatment),
         group=as.factor(group))
# df %>% 
#   filter(treatment=="Untreated") %>% 
#   filter(time==0) %>% 
#   mutate(conc=)
```

You need to duplicate each person's untreated hand count and label it as treatment to allow for the same statrting value
```{r}

df %>% 
  filter(treatment=="Untreated") %>% 
  filter(time==0) %>% 
  mutate()
```


## Including Plots

Let's have a quick look at each person's individual trajectory. Note that individuals are in both "treated" and "untreated" groups because only one hand is washed with the other one being thte control

```{r Plot the data - Individual, echo=FALSE}
df %>% 
  ggplot()+
  geom_point(aes(x=time,y=conc,colour=id %>% as.factor()),size=1.5)+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x)
 )+
  annotation_logticks(sides = "l")  +
  scale_color_brewer(palette = "Set1")+
  xlab("Hours after cleaning")+
  ylab("Concentration [CFU/cm2]")+
  labs(col="Individual participant")+
  hrbrthemes::theme_ipsum()
  
  
```

Here we plot the points and superimpose a meand value plus 95% confidence interval

```{r Plot the data - Aggregate, echo=FALSE}
df %>% 
  ggplot()+
  geom_point(aes(x=time,y=conc),colour="gray",alpha=0.5,size=1.5)+
  # stat_summary(aes(x=time,y=conc,colour=treatment),fun.data = "mean_cl_boot", size = 0.8)+
  stat_summary(aes(x=time,y=conc,colour=treatment),fun.data = "mean_cl_boot", size = 1.01)+
  # scale_y_continuous(trans = "log10")+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x)
 )+
  annotation_logticks(sides = "l")  +
  scale_color_brewer(palette = "Set1")+
  xlab("Hours after cleaning")+
  ylab("Colony forming units [CFU]")+
  labs(col="Group")+
  hrbrthemes::theme_ipsum()
  
  
```

# Simple regression model


```{r}
df %>% 
  ggplot()+
  geom_point(aes(x=time,y=conc),colour="gray",alpha=0.5,size=1.5)+
  stat_summary(aes(x=time,y=conc,colour=treatment),fun.data = "mean_cl_boot", size = 1.01)+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x)
 )+
  annotation_logticks(sides = "l")  +
  scale_color_brewer(palette = "Set1")+
  geom_smooth(method="lm",aes(x=time,y=conc,colour=treatment))+
  xlab("Hours after cleaning")+
  ylab("Colony forming units [CFU]")+
  labs(col="Group")+
  hrbrthemes::theme_ipsum()
```


#Run a regression model to estimate the effect of time

Use the bayesian regression package BRMS and plots using bayesplot "group+time+intervention + (time+intervention | participant)"
```{r}

fit_m <- brm(log10(conc) ~ 1+group+time+treatment+ (time+treatment | id), 
                              data   = df, 
                              warmup = 1000, 
                              iter   = 3000, 
                              chains = 4, 
                              inits  = "random",
                              cores  = 4)
summary(fit_m)

plot(fit_m)

```


Using bayesplot package we can create some nice plots which investigate the fitted model

```{r}
color_scheme_set("red")
ppc_dens_overlay(y = log10(fit_m$data$conc),
                 yrep = posterior_predict(fit_m, draws = 10))
```

